### 2022-12-19 intro.Rnw
last polish of intro?

\begin{wrapfigure}{r}{0.29\textwidth}
  \centering
  \includegraphics[width = 0.2\textwidth]{images/xkcdCrop.png}
  \caption*{Replication studies as illustrated by Randall Munroe
    (\url{https://xkcd.com/242}).}
\end{wrapfigure}

@@

% How can we know if a finding from a scientific study is really true? For
% example, how can we know if the protective effect of a vaccine found in a study
% is real? The answer to such questions is of high importance to scientists,
% decision makers, and the general population. However, because any study result
% comes with uncertainty we can never know for sure.
Being able to trust scientific findings is essential for making evidence-based
decisions. But how can we know whether a scientific finding is trustworthy? For
instance, how can we know whether the protective effect of a vaccine found in a
study is indeed existent, or whether it is just a spurious finding? One way to
answer this question is through the use of \emph{replication studies}.

A replication study (or, simply, replication) involves repeating the original
study as closely as possible but with new subjects. If the replication yields
similar results to the original study, this increases our confidence in the
original finding. However, if the replication study yields conflicting results,
this lowers our confidence in the original finding. %  As such, replication studies
% are an essential part of the scientific method. They provide a means for
% substantiating genuine research findings and refuting findings which occurred
% merely by chance.
The concept of replicability has long been a central part of the scientific
method. For instance, \citet[p. 13]{Fisher1935} noted ``\textit{no isolated
  experiment, however significant in itself, can suffice for the experimental
  demonstration of any natural phenomenon}''. For this reason, ``successful''
replication is typically a requirement for acceptance of newly proposed
scientific theories (\eg{} a new physical model). Yet, replication studies are
not only important for researchers, but also for decision makers and the general
population, as the implementation of policies based on scientific findings may
also depend on whether these findings can be replicated. For instance, the
United States food and drug administration requires ``\textit{at least two
  adequate and well-controlled studies, each convincing on its own, to establish
  effectiveness}'' \citep[p. 3]{FDA1998} for market approval of a drug. The
results from replication studies may hence have real-world consequences. It is
therefore important that the methodological aspects of replication studies, such
as their statistical design and analysis, are well understood.

Despite their importance, the traditional academic system has made it
unattractive for researchers to conduct replication studies; publications,
citations, and grant money are typically easier to acquire by conducting novel
research. This is because until recently, most scientific journals hardly
published any replication studies as the emphasis was on ``innovation'' rather
than on ``repetition'' \citep{Makel2012, Martin2017, Coiera2021}. As a result,
researchers often had no other choice than to focus their efforts on producing
novel and eye-catching research results in order to build successful careers
\citep{Binswanger2014, Moher2018}.

@@

In the midst of this crisis, % , particularly from the hard-hit
% field of psychology,
various reforms were implemented to prevent an escalation of the situation % and
% save the reputation of science
\citep[for an overview see \eg{}][]{Munafo2017}. For instance, some journals
adopted the ``registered report'' format \citep{Chambers2021} in which a study
proposal is peer reviewed \emph{before} the study is conducted, and which, if
accepted, gives provisional publication acceptance regardless of the study
outcome. Similarly, digital infrastructure platforms, such as Zenodo
(\url{https://zenodo.org}) or the Open Science Framework (\url{https://osf.io}),
were created to facilitate preregistration, preprints, code and data sharing,
all of which have substantially increased over the last decade
\citep{Kidwell2016, Nosek2018, Rawlinson2019}. The practice of conducting
replication studies has also gained popularity. Several academic journals and
funders are now explicitly promoting replication research \citep{NWO2016,
  NSL2018, Nature2022}. Figure~\ref{fig:scholar} (left) illustrates this trend
by the yearly number of results for the search term ``replication study''
obtained from the search engine Web of Science. The numbers have been rapidly
growing, especially since the mid-2000s. This trend could possibly be explained
by the fact that more research results are published each year but also when the
numbers are normalized by the number of results for the search term ``study'',
the increasing trend is still visible.

@@

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. % Effect
% estimates, standard error, confidence intervals, and $p$-values
All of these are routinely reported in research articles, and if one of them is
missing they can typically be calculated from the others. Using summary
statistics is also often the only possible way for conducting the replication
success analysis as the raw data from the original study may not be available to
the researchers conducting the replication study. The most important statistics
are the effect estimates $\that_{o}$ and $\that_{r}$. They provide an estimate
of the underlying effect size $\theta$ which quantifies the true effect or
association of an intervention/exposure with an outcome variable. Typical effect
sizes are mean differences and correlations (for continuous outcomes), odds
ratios (for binary outcomes), or hazard ratios (for time-to-event outcomes).
Depending on the type of effect size, a transformation might be required so that
the assumption of approximately normally distributed effect estimates around the
unknown effect size $\theta$ (for large enough sample sizes) is justified. This
could be, for instance, the Fisher $z$-transformation for correlations or the
log-transformation for odds/hazard ratios \citep[chapter 11]{Cooper2019}. The
associated standard errors $\sigma_{o}$ and $\sigma_{r}$ represent the
statistical uncertainty of the estimates. Typically, the available standard
errors are only estimates of the true standard errors. However, in most parts of
this thesis it will be assumed that $\sigma_{o}$ and $\sigma_{r}$ correspond to
the true standard errors. This is the same assumption as in ordinary
fixed-effect/random-effects meta-analysis, and typically reasonable if the
sample size of the studies is not too small. Under the assumption of
(asymptotic) normality, confidence intervals for $\theta$ and $p$-values for
testing the null hypothesis $H_{0} \colon \theta = 0$ can be computed as shown
in Table~\ref{tab:meta}.

\begin{table}[!htb]
  \centering
  \caption{Statistical criteria for assessing replication success which have
    been used in practice \citep{Opensc2015, Camerer2016, Camerer2018, Cova2018,
      Errington2021}.}
  \label{tab:rsdef}
  \begin{tabular}{p{0.21\textwidth}p{0.7\textwidth}}
    \toprule
    Criterion type & Description \\
    \midrule
    Significance & The original and replication $p$-values are smaller than
                   a threshold $\alpha$ and their effect estimates show
                   the same direction
                   (\mbox{$p_{o} < \alpha$}, \mbox{$p_{r} < \alpha$}, and
                   \mbox{$\sign(\that_{o}) = \sign(\that_{r})$}). Usually $\alpha = 5\%$.
    \\
                   & \\
    Meta-analytic significance & The meta-analytic $p$-value is smaller than a threshold
                                 $\alpha$ (\mbox{$p_{m} = 2\{1 - \Phi(|\that_{m}|/\sigma_{m})\} < \alpha$})
                                 where \mbox{$\that_{m} =
                                 (\that_{o}/\sigma_{o}^{2} + \that_{r}/\sigma^{2}_{r})\sigma^{2}_{m}$}
                                 is the  pooled effect estimate with standard error
                                 \mbox{$\sigma_{m} = (1/\sigma^{2}_{o} + 1/\sigma^{2}_{r})^{-1/2}$}.
                                 There are no established conventions regarding the level $\alpha$.
    \\
                   & \\
    Relative effect size & The effect estimate of the replication study goes in the same
                           direction as the original one and is at least $d_{\text{min}}$ times as large (\mbox{$d = \that_{r}/\that_{o} \geq d_{\text{min}}$}).
                           Usually $d_{\text{min}} = 1$. \\
                   & \\
    Confidence interval & The replication effect estimate is contained in the
                          $(1 - \alpha)$ original confidence interval
                          ($\that_{r} \in [\that_{o} - z_{\alpha/2}\sigma_{o},
                          \that_{o} + z_{\alpha/2}\sigma_{o}]$).
                          Sometimes, also defined as the original effect estimate is
                          contained in the $(1 - \alpha)$ replication confidence interval
                          ($\that_{o} \in [\that_{r} - z_{\alpha/2}\sigma_{r},
                          \that_{r} + z_{\alpha/2}\sigma_{r}]$).
                          Usually $\alpha = 5\%$. \\
                   & \\
    Prediction interval ($Q$-test) & The replication effect estimate is contained in its
                                     $(1 - \alpha)$ prediction interval
                                     ($\that_{r} \in [\that_{o} - z_{\alpha/2}\surd(\sigma_{o}^{2}
                                     + \sigma_{r}^{2}), \that_{o} + z_{\alpha/2}\surd(\sigma_{o}^{2}
                                     + \sigma_{r}^{2})]$). Usually $\alpha = 5\%$.
                                     This criterion is equivalent to $p_{Q} \geq \alpha$ where
                                     $p_{Q}$ is the $p$-value from the meta-analytic $Q$-test
                                     $p_{Q} = 2\{1 - \Phi(\surd Q)\}$ with $Q$-statistic
                                     $Q = \sum_{i \in \{o, r\}} (\that_{i} - \that_{m})^{2}/\sigma_{i}^{2} = (\that_{o}- \that_{r})^{2}/(\sigma^{2}_{o} + \sigma^{2}_{r})$.\\
    \bottomrule
  \end{tabular}
\end{table}

@@

The remaining criteria in Table~\ref{tab:rsdef} put more emphasis on
compatibility in effect size between the original and the replication study. For
example, the relative effect estimate $d = \that_{r}/\that_{o}$ quantifies how
much the magnitude of the replication effect estimate changed compared to the
original, and the smaller $d$ is the smaller the degree of replication success.
Some projects also report a confidence interval for $d$ \citep{Camerer2016,
  Camerer2018}, whereas others ignore its uncertainty and make a binary cut at
$d_{\text{min}} = 1$ to define replication success \citep{Errington2021}. The
criteria based on confidence and prediction intervals define effect size
compatibility on an absolute scale. However, the criterion based on the original
confidence interval ignores the uncertainty from the replication, whereas the
criterion based on the replication confidence interval ignores the uncertainty
from the original study. The prediction interval criterion takes into account
both sources of uncertainty \citep{Patil2016}. Yet, declaring ``replication
success'' based on the prediction interval may be logically questionable due to
its connection to the meta-analytic $Q$-test.
% as it may also be seen as an instance of the ``absence of evidence fallacy''
% \citep{Altman1995};
That is, if the $p$-value from the $Q$-test is larger than $\alpha$ this is
equivalent to the replication effect estimate $\that_{r}$ being contained in its
$(1- \alpha)$ prediction interval. The null hypothesis of this test is defined
that the underlying effect sizes of original and replication study are the same
($H_{0} \colon \theta_{o} = \theta_{r}$), so a rejection of this null hypothesis
corresponds to demonstrating replication failure and not replication success.
Interpreting a failure to reject the null hypothesis as evidence for it is again
an instance of the ``absence of evidence fallacy''
\citep{Hedges2019}. % To establish replication success a
% test based on the null hypothesis $H_{0} \colon \theta_{o} \neq \theta_{r}$
% would be required \citep{Hedges2019}.
As in the case of defining replication success by simultaneous non-significance
of original and replication study, the mismatch of the null hypothesis of the
prediction interval criterion results in the undesirable property that
``replication success'' can almost always be achieved if the sample size of the
studies is small enough, because the prediction interval becomes wider with
larger standard errors.

@@

\begin{figure}[!htb]
<< "RPCB-graph", fig.height = 3.5 >>=
za <- qnorm(p = 0.975)
rpcb <- read.csv("data/rpcb.csv") |>
    mutate(id = paste0("(", paper, ", ", experiment, ", ", effect, ")"),
           zo = smdo/so,
           zr = smdr/sr,
           c = so^2/sr^2,
           d = smdr/smdo,
           po2 = 2*(1 - pnorm(q = abs(zo))),
           pr2 = 2*(1 - pnorm(q = abs(zr))),
           sm = 1/sqrt(1/so^2 + 1/sr^2),
           smdm = (smdo/so^2 + smdr/sr^2)*sm^2,
           pm2 = 2*(1 - pnorm(q = abs(smdm/sm))),
           Q = (smdo - smdr)^2/(so^2 + sr^2),
           pQ = pchisq(q = Q, df = 1, lower.tail = FALSE),
           rsig = factor(pr2 <= 0.05, levels = c(FALSE, TRUE),
                         labels = c("italic(p[r]) > 0.05",
                                    "italic(p[r]) <= 0.05")))
## compute sceptical p-value (TODO fix problems with NA)
ind <- !is.na(rpcb$zo) & !is.na(rpcb$zr) & !is.na(rpcb$c)
rpcb$ps1 <- NA
rpcb$ps1[ind] <- pSceptical(zo = rpcb$zo[ind], zr = rpcb$zr[ind], c = rpcb$c[ind],
                            alternative = "one.sided", type = "nominal")

zoom <- c(-2, 12)
alphazoom <- 0.6
ltyzoom <- 3
rpcbWithoutNA <- na.omit(rpcb)
plotA <- ggplot(data = rpcbWithoutNA, aes(x = smdo, y = smdr)) +
    annotate(geom = "segment", x = zoom[1], xend = zoom[2],
             y = zoom[1], yend = zoom[1], alpha = alphazoom,
             lty = ltyzoom) +
    annotate(geom = "segment", x = zoom[1], xend = zoom[2],
             y = zoom[2], yend = zoom[2], alpha = alphazoom,
             lty = ltyzoom) +
    annotate(geom = "segment", x = zoom[1], xend = zoom[1],
             y = zoom[1], yend = zoom[2], alpha = alphazoom,
             lty = ltyzoom) +
    annotate(geom = "segment", x = zoom[2], xend = zoom[2],
             y = zoom[1], yend = zoom[2], alpha = alphazoom,
             lty = ltyzoom) +
    geom_abline(intercept = 0, slope = 1, alpha = 0.3) +
    geom_hline(yintercept = 0, lty = 2, alpha = 0.9, size = 0.4) +
    geom_vline(xintercept = 0, lty = 2, alpha = 0.9, size = 0.4) +
    geom_point(aes(fill = rsig), shape = 21, alpha = 0.8) +
    geom_rug(aes(color = rsig), size = 0.3, show.legend = FALSE) +
    labs(x = "Original effect estimate (SMD)",
         y = "Replication effect estimate (SMD)",
         fill = "") +
    coord_fixed(xlim = c(-5, 30), ylim = c(-5, 30)) +
    scale_fill_discrete(labels = scales::parse_format()) +
    theme_bw() +
    theme(panel.grid = element_blank(),
          legend.position = c(0.5, 0.835))

plotB <- ggplot(data = rpcbWithoutNA, aes(x = smdo, y = smdr)) +
    geom_abline(intercept = 0, slope = 1, alpha = 0.3) +
    geom_hline(yintercept = 0, lty = 2, alpha = 0.9, size = 0.4) +
    geom_vline(xintercept = 0, lty = 2, alpha = 0.9, size = 0.4) +
    geom_point(aes(fill = rsig), shape = 21, alpha = 0.8,
               show.legend = FALSE) +
    geom_rug(aes(color = rsig), size = 0.3, show.legend = FALSE) +
    labs(x = "Original effect estimate (SMD)",
         y = "Replication effect estimate (SMD)",
         fill = "") +
    coord_fixed(xlim = zoom, ylim = zoom) +
    theme_bw() +
    theme(panel.grid = element_blank())
ggpubr::ggarrange(plotA, plotB, ncol = 2, align = "v")
@
\caption{Results for 132 effects from the ``Reproducibility Project: Cancer
  Biology'' \citep{Errington2021} for which effect estimates and standard errors
  are available on standardized mean difference (SMD) scale. The right plot
  shows a zoomed-in view of the dotted area in the left plot. The
  \textit{p}-values are recomputed using a normal approximation. Two study pairs
  with original effect estimate larger than 80 are not shown.}
\label{fig:rpcb}
\end{figure}

@@

\begin{figure}[!htb]
<< "RPCB-table1", results = "asis", fig.height = 6 >>=
## show number of successes
alpha <- 0.05
dmin <- 1
rpcbUpset<- rpcb |>
    filter(!is.na(po2), !is.na(pr2), !is.na(smdo), !is.na(smdr),
           !is.na(so), !is.na(sr)) |>
    mutate("Significance" = as.numeric(po2 < alpha & pr2 < alpha & sign(smdo) == sign(smdr)),
           "Meta analytic \nsignificance" = as.numeric(pm2 < alpha),
           "Relative \neffect size" = as.numeric(d >= dmin),
           "Prediction \ninterval" = as.numeric(pQ > alpha))
nsig <- sum(rpcbUpset[,"Significance"], na.rm = TRUE)
nma <- sum(rpcbUpset[,"Meta analytic \nsignificance"], na.rm = TRUE)
nres <- sum(rpcbUpset[,"Relative \neffect size"], na.rm = TRUE)
npi <- sum(rpcbUpset[,"Prediction \ninterval"], na.rm = TRUE)

@@

\begin{landscape}
\begin{table}
  \centering
  \caption{Subset of results from the ``Reproducibility Project: Cancer
    Biology'' \citep{Errington2021}. (P, X, E) denotes effect number E from
    experiment number X, from original paper number P. Shown are original
    $\that_{o}$, replication $\that_{r}$, and pooled effect estimate $\that_{m}$
    with 95\% confidence intervals, variance ratio
    $c = \sigma^{2}_{o}/\sigma^{2}_{r}$, relative effect estimate
    $d = \that_{r}/\that_{o}$, original \textit{p}-value $p_{o}$, replication
    \textit{p}-value $p_{r}$, meta-analytic \textit{p}-value $p_{m}$, $Q$-test
    \textit{p}-value $p_{Q}$, and one-sided sceptical \textit{p}-value
    $p_{\mathrm{S}}$. A replication success (green) threshold of $\alpha = 5\%$
    is used for two-sided \textit{p}-values, $\alpha = 2.5\%$ for one-sided
    $p$-values, and $d_{\text{min}} = 1$ for the relative effect estimate.}
  \label{tab:RPCB}
\resizebox{1.02\linewidth}{!}{%
<< "RPCB-table2", results = "asis" >>=
## function to format CI
ciFormat. <- function(lower, upper, dollar = FALSE, digits = 2) {
    if (dollar) x <- "$" else x <- ""
    paste0("(", x, format(round(lower, digits), nsmall = digits), x,
           " to ", x, format(round(upper, digits), nsmall = digits), x, ")")
}
ciFormat <- Vectorize(FUN = ciFormat.)

>>>>

\begin{wrapfigure}{r}{0.285\textwidth}
  \centering
  \includegraphics[width = 0.2\textwidth]{images/xkcdCrop.png}
  \caption*{Replication studies as illustrated by Randall Munroe
    (\url{https://xkcd.com/242}).}
\end{wrapfigure}

@@

% How can we know if a finding from a scientific study is really true? For
% example, how can we know if the protective effect of a vaccine found in a study
% is real? The answer to such questions is of high importance to scientists,
% decision makers, and the general population. However, because any study result
% comes with uncertainty we can never know for sure.
Being able to trust scientific findings is essential for making evidence-based
decisions. But how can we know whether a scientific finding is trustworthy? For
instance, how can we know whether the protective effect of a vaccine found in a
study is indeed existent? One way to answer this question is through the use of
\emph{replication studies}.

A replication study (or simply replication) involves repeating the original
study as closely as possible but with new subjects. If the replication yields
similar results to the original study, this increases our confidence in the
original finding. However, if the replication study yields conflicting results,
this lowers our confidence in the original finding. %  As such, replication studies
% are an essential part of the scientific method. They provide a means for
% substantiating genuine research findings and refuting findings which occurred
% merely by chance.
The concept of replicability has long been a central part of the scientific
method. For instance, \citet[p. 13]{Fisher1935} noted ``\textit{no isolated
  experiment, however significant in itself, can suffice for the experimental
  demonstration of any natural phenomenon}''. For this reason, ``successful''
replication is typically a requirement for acceptance of newly proposed
scientific theories (\eg{} a new physical model). Yet, replication studies are
not only important for researchers, but also for decision makers and the general
population, as the implementation of policies based on scientific findings may
also depend on whether these findings can be replicated. For instance, for a
drug to be approved to the market, the United States food and drug
administration requires ``\textit{at least two adequate and well-controlled
  studies, each convincing on its own, to establish effectiveness}'' \citep[p.
3]{FDA1998}. The results from replication studies may hence have real-world
consequences. It is therefore important that the methodological aspects of
replication studies, such as their statistical design and analysis, are well
understood.

Despite their importance, the traditional academic system has made it
unattractive for researchers to conduct replication studies as publications,
citations, and grant money are typically easier to acquire by conducting novel
research. This is because until recently, most scientific journals hardly
published any replication studies since the emphasis was on ``innovation''
rather than on ``repetition'' \citep{Makel2012, Martin2017, Coiera2021}. As a
result, researchers often had no other choice than to focus their efforts on
producing novel and eye-catching research results in order to build successful
careers \citep{Binswanger2014, Moher2018}.

@@

In the midst of this crisis, various reforms were implemented to prevent an
escalation of the situation \citep[for an overview see \eg{}][]{Munafo2017}. For
instance, some journals adopted the ``registered report'' format
\citep{Chambers2021} in which a study proposal is peer reviewed \emph{before}
the study is conducted, and which, if accepted, gives provisional publication
acceptance regardless of the study outcome. Similarly, digital infrastructure
platforms, such as Zenodo (\url{https://zenodo.org}) or the Open Science
Framework (\url{https://osf.io}), were created to facilitate pre-registration,
preprints, code and data sharing, all of which have substantially increased over
the last decade \citep{Kidwell2016, Nosek2018, Rawlinson2019}. The practice of
conducting replication studies has also gained popularity. Several academic
journals and funders are now explicitly promoting replication research
\citep{NWO2016, NSL2018, Nature2022}. Figure~\ref{fig:scholar} (left)
illustrates this trend by the yearly number of results for the search term
``replication study'' obtained from the search engine Web of Science. The
numbers have been rapidly growing, especially since the mid-2000s. This trend
could possibly be explained by the fact that more research results are published
each year but also when the numbers are normalized by the number of results for
the search term ``study'', the increasing trend is still visible.

@@

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. All of these
are routinely reported in research articles, and if one of them is missing they
can typically be calculated from the others. Using summary statistics is also
often the only possible way for conducting the replication success analysis as
the raw data from the original study may not be available to the researchers
conducting the replication study. The most important statistics are the effect
estimates $\that_{o}$ and $\that_{r}$. They provide an estimate of the
underlying effect size $\theta$ which quantifies the true effect or association
of an intervention/exposure with an outcome variable. Typical effect sizes are
mean differences and correlations (for continuous outcomes), odds ratios (for
binary outcomes), or hazard ratios (for time-to-event outcomes). Depending on
the type of effect size, a transformation might be required so that the
assumption of approximately normally distributed effect estimates around the
unknown effect size $\theta$ (for large enough sample sizes) is justified. This
could be, for instance, the Fisher $z$-transformation for correlations or the
log-transformation for odds/hazard ratios \citep[chapter 11]{Cooper2019}. The
associated standard errors $\sigma_{o}$ and $\sigma_{r}$ represent the
statistical uncertainty of the estimates. Typically, the available standard
errors are only estimates of the true standard errors. However, in most parts of
this thesis it will be assumed that $\sigma_{o}$ and $\sigma_{r}$ correspond to
the true standard errors. This is the same assumption as in ordinary
fixed-effect/random-effects meta-analysis, and typically reasonable if the
sample size of the studies is not too small. Under the assumption of
(asymptotic) normality, confidence intervals for $\theta$ and $p$-values for
testing the null hypothesis $H_{0} \colon \theta = 0$ can be computed as shown
in Table~\ref{tab:meta}.

\begin{table}[!htb]
  \centering
  \caption{Statistical criteria for assessing replication success which have
    been used in practice \citep{Opensc2015, Camerer2016, Camerer2018, Cova2018,
      Errington2021}.}
  \label{tab:rsdef}
  \begin{tabular}{p{0.21\textwidth}p{0.7\textwidth}}
    \toprule
    Criterion type & Description \\
    \midrule
    Significance & \sloppy The original and replication $p$-values are smaller than
                   a threshold $\alpha$ and their effect estimates show
                   the same direction
                   (\mbox{$p_{o} < \alpha$}, \mbox{$p_{r} < \alpha$}, and
                   \mbox{$\sign(\that_{o}) = \sign(\that_{r})$}). Usually $\alpha = 5\%$.
    \\
                   & \\
    Meta-analytic significance & The meta-analytic $p$-value is smaller than a threshold
                                 $\alpha$ (\mbox{$p_{m} = 2\{1 - \Phi(|\that_{m}|/\sigma_{m})\} < \alpha$})
                                 where \mbox{$\that_{m} =
                                 (\that_{o}/\sigma_{o}^{2} + \that_{r}/\sigma^{2}_{r})\sigma^{2}_{m}$}
                                 is the  pooled effect estimate with standard error
                                 \mbox{$\sigma_{m} = (1/\sigma^{2}_{o} + 1/\sigma^{2}_{r})^{-1/2}$}.
                                 There are no established conventions regarding the level $\alpha$.
    \\
                   & \\
    Relative effect size & The effect estimate of the replication study goes in the same
                           direction as the original one and is at least $d_{\text{min}}$ times as large (\mbox{$d = \that_{r}/\that_{o} \geq d_{\text{min}}$}).
                           Usually $d_{\text{min}} = 1$. \\
                   & \\
    Confidence interval & \sloppy The replication effect estimate is contained in the
                          \mbox{$(1 - \alpha)\times100\%$} original confidence interval
                          ($\that_{r} \in [\that_{o} - z_{\alpha/2}\sigma_{o},
                          \that_{o} + z_{\alpha/2}\sigma_{o}]$).
                          Sometimes, also defined as the original effect estimate is
                          contained in the \mbox{$(1 - \alpha)\times100\%$} replication
                          confidence interval
                          (\mbox{$\that_{o} \in [\that_{r} - z_{\alpha/2}\sigma_{r},
                          \that_{r} + z_{\alpha/2}\sigma_{r}]$}).
                          Usually $\alpha = 5\%$. \\
                   & \\
    Prediction interval ($Q$-test) & The replication effect estimate is contained in its
                                     $(1 - \alpha)\times100\%$ prediction interval
                                     \mbox{($\that_{r} \in [\that_{o} - z_{\alpha/2}\surd(\sigma_{o}^{2}
                                     + \sigma_{r}^{2}), \that_{o} + z_{\alpha/2}\surd(\sigma_{o}^{2}
                                     + \sigma_{r}^{2})]$)}. Usually $\alpha = 5\%$.
                                     This criterion is equivalent to $p_{Q} \geq \alpha$ where
                                     $p_{Q}$ is the $p$-value from the meta-analytic $Q$-test
                                     $p_{Q} = 2\{1 - \Phi(\surd Q)\}$ with $Q$-statistic
                                     $Q = \sum_{i \in \{o, r\}} (\that_{i} - \that_{m})^{2}/\sigma_{i}^{2} = (\that_{o}- \that_{r})^{2}/(\sigma^{2}_{o} + \sigma^{2}_{r})$.\\
    \bottomrule
  \end{tabular}
\end{table}

@@

The remaining criteria in Table~\ref{tab:rsdef} put more emphasis on
compatibility in effect size between the original and the replication study. For
example, the relative effect estimate $d = \that_{r}/\that_{o}$ quantifies how
much the magnitude of the replication effect estimate changed compared to the
original, and the smaller $d$ is the smaller the degree of replication success.
Some projects also report a confidence interval for $d$ \citep{Camerer2016,
  Camerer2018}, whereas others ignore its uncertainty and make a binary cut at
$d_{\text{min}} = 1$ to define replication success \citep{Errington2021}. The
criteria based on confidence and prediction intervals define effect size
compatibility on an absolute scale. However, the criterion based on the original
confidence interval ignores the uncertainty from the replication, whereas the
criterion based on the replication confidence interval ignores the uncertainty
from the original study. The prediction interval criterion takes into account
both sources of uncertainty \citep{Patil2016}. Yet, declaring ``replication
success'' based on the prediction interval may be logically questionable due to
its connection to the meta-analytic \mbox{$Q$-test}. That is, if the $p$-value
from the $Q$-test is larger than $\alpha$ this is equivalent to the replication
effect estimate $\that_{r}$ being contained in its $(1- \alpha)\times 100\%$
prediction interval. The null hypothesis of this test is defined that the
underlying effect sizes of original and replication study are the same
($H_{0} \colon \theta_{o} = \theta_{r}$), so a rejection of this null hypothesis
corresponds to demonstrating replication failure and not replication success.
Interpreting a failure to reject the null hypothesis as evidence for it is again
an instance of the ``absence of evidence fallacy''
\citep{Hedges2019}. % To establish replication success a
% test based on the null hypothesis $H_{0} \colon \theta_{o} \neq \theta_{r}$
% would be required \citep{Hedges2019}.
As in the case of defining replication success by simultaneous non-significance
of original and replication study, the mismatch of the null hypothesis of the
prediction interval criterion results in the undesirable property that
``replication success'' can almost always be achieved if the sample size of the
studies is small enough, because the prediction interval becomes wider with
larger standard errors.

@@

\begin{figure}[!htb]
<< "RPCB-graph", fig.height = 3.75 >>=
za <- qnorm(p = 0.975)
rpcb <- read.csv("data/rpcb.csv") |>
    mutate(id = paste0("(", paper, ", ", experiment, ", ", effect, ")"),
           zo = smdo/so,
           zr = smdr/sr,
           c = so^2/sr^2,
           d = smdr/smdo,
           po2 = 2*(1 - pnorm(q = abs(zo))),
           pr2 = 2*(1 - pnorm(q = abs(zr))),
           sm = 1/sqrt(1/so^2 + 1/sr^2),
           smdm = (smdo/so^2 + smdr/sr^2)*sm^2,
           pm2 = 2*(1 - pnorm(q = abs(smdm/sm))),
           Q = (smdo - smdr)^2/(so^2 + sr^2),
           pQ = pchisq(q = Q, df = 1, lower.tail = FALSE),
           rsig = factor(pr2 <= 0.05, levels = c(FALSE, TRUE),
                         labels = c("italic(p[r]) > 0.05",
                                    "italic(p[r]) <= 0.05")))
## compute sceptical p-value (TODO fix problems with NA)
ind <- !is.na(rpcb$zo) & !is.na(rpcb$zr) & !is.na(rpcb$c)
rpcb$ps1 <- NA
rpcb$ps1[ind] <- pSceptical(zo = rpcb$zo[ind], zr = rpcb$zr[ind], c = rpcb$c[ind],
                            alternative = "one.sided", type = "nominal")

zoom <- c(-2, 12)
alphazoom <- 0.6
ltyzoom <- 3
rpcbWithoutNA <- na.omit(rpcb)
cols <- c("#1A1A1A80", "#1B9E77")
names(cols) <- levels(rpcbWithoutNA$rsig)
plotA <- ggplot(data = rpcbWithoutNA, aes(x = smdo, y = smdr)) +
    annotate(geom = "segment", x = zoom[1], xend = zoom[2],
             y = zoom[1], yend = zoom[1], alpha = alphazoom,
             lty = ltyzoom) +
    annotate(geom = "segment", x = zoom[1], xend = zoom[2],
             y = zoom[2], yend = zoom[2], alpha = alphazoom,
             lty = ltyzoom) +
    annotate(geom = "segment", x = zoom[1], xend = zoom[1],
             y = zoom[1], yend = zoom[2], alpha = alphazoom,
             lty = ltyzoom) +
    annotate(geom = "segment", x = zoom[2], xend = zoom[2],
             y = zoom[1], yend = zoom[2], alpha = alphazoom,
             lty = ltyzoom) +
    geom_abline(intercept = 0, slope = 1, alpha = 0.3) +
    geom_hline(yintercept = 0, lty = 2, alpha = 0.9, size = 0.4) +
    geom_vline(xintercept = 0, lty = 2, alpha = 0.9, size = 0.4) +
    geom_point(aes(fill = rsig), shape = 21, alpha = 0.9) +
    geom_rug(aes(color = rsig), size = 0.3, show.legend = FALSE) +
    labs(x = "Original effect estimate (SMD)",
         y = "Replication effect estimate (SMD)",
         fill = "") +
    coord_fixed(xlim = c(-5, 30), ylim = c(-5, 30)) +
    scale_fill_manual(values = cols, labels = scales::parse_format()) +
    scale_color_manual(values = cols) +
    theme_bw() +
    theme(panel.grid = element_blank(),
          legend.position = c(0.5, 0.835))

plotB <- ggplot(data = rpcbWithoutNA, aes(x = smdo, y = smdr)) +
    geom_abline(intercept = 0, slope = 1, alpha = 0.3) +
    geom_hline(yintercept = 0, lty = 2, alpha = 0.9, size = 0.4) +
    geom_vline(xintercept = 0, lty = 2, alpha = 0.9, size = 0.4) +
    geom_point(aes(fill = rsig), shape = 21, alpha = 0.9,
               show.legend = FALSE) +
    geom_rug(aes(color = rsig), size = 0.3, show.legend = FALSE) +
    labs(x = "Original effect estimate (SMD)",
         y = "Replication effect estimate (SMD)",
         fill = "") +
    coord_fixed(xlim = zoom, ylim = zoom) +
    scale_fill_manual(values = cols, labels = scales::parse_format()) +
    scale_color_manual(values = cols) +
    theme_bw() +
    theme(panel.grid = element_blank())
ggpubr::ggarrange(plotA, plotB, ncol = 2, align = "v")
@
\caption{Results for 132 effects from the ``Reproducibility Project: Cancer
  Biology'' \citep{Errington2021} for which effect estimates and standard errors
  are available on standardized mean difference (SMD) scale. The right plot
  shows a zoomed-in view of the dotted area in the left plot. The
  \textit{p}-values are recomputed using a normal approximation. Two study pairs
  with original effect estimate larger than 80 are not shown.}
\label{fig:rpcb}
\end{figure}

@@

\begin{figure}[!htb]
<< "RPCB-table1", results = "asis", fig.height = 6.5 >>=
## show number of successes
alpha <- 0.05
dmin <- 1
rpcbUpset<- rpcb |>
    filter(!is.na(po2), !is.na(pr2), !is.na(smdo), !is.na(smdr),
           !is.na(so), !is.na(sr)) |>
    mutate("Significance" = as.numeric(po2 < alpha & pr2 < alpha & sign(smdo) == sign(smdr)),
           "Meta analytic \nsignificance" = as.numeric(pm2 < alpha),
           "Relative \neffect size" = as.numeric(d >= dmin),
           "Prediction \ninterval" = as.numeric(pQ > alpha))
nsig <- sum(rpcbUpset[,"Significance"], na.rm = TRUE)
nma <- sum(rpcbUpset[,"Meta analytic \nsignificance"], na.rm = TRUE)
nres <- sum(rpcbUpset[,"Relative \neffect size"], na.rm = TRUE)
npi <- sum(rpcbUpset[,"Prediction \ninterval"], na.rm = TRUE)

@@

\begin{landscape}
\begin{table}
  \centering
  \caption{Subset of results from the ``Reproducibility Project: Cancer
    Biology'' \citep{Errington2021}. (P, X, E) denotes effect number E from
    experiment number X, from original paper number P. Shown are original
    $\that_{o}$, replication $\that_{r}$, and pooled effect estimate $\that_{m}$
    with 95\% confidence intervals, variance ratio
    $c = \sigma^{2}_{o}/\sigma^{2}_{r}$, relative effect estimate
    $d = \that_{r}/\that_{o}$, original \textit{p}-value $p_{o}$, replication
    \textit{p}-value $p_{r}$, meta-analytic \textit{p}-value $p_{m}$, $Q$-test
    \textit{p}-value $p_{Q}$, and one-sided sceptical \textit{p}-value
    $p_{\mathrm{S}}$. A replication success (green-bold) threshold of $\alpha = 5\%$
    is used for two-sided \textit{p}-values, $\alpha = 2.5\%$ for one-sided
    $p$-values, and $d_{\text{min}} = 1$ for the relative effect estimate.}
  \label{tab:RPCB}
\resizebox{1.02\linewidth}{!}{%
<< "RPCB-table2", results = "asis" >>=
## function to format CI
ciFormat. <- function(lower, upper, dollar = FALSE, digits = 2) {
    if (dollar) x <- "$" else x <- ""
    paste0("(", x, format(round(lower, digits), nsmall = digits), x,
           " to ", x, format(round(upper, digits), nsmall = digits), x, ")")
}
ciFormat <- Vectorize(FUN = ciFormat.)
